---
title: "Tidy Tuesday 35"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
library(pell)
library(bbplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(gganimate)
library(tidyr)

```

## Loading the dataset and some minor cleaning

```{r}
dat <- pell %>% janitor::clean_names() %>%
  mutate(year = as.Date(ISOdate(year, 1, 1)))

```

## Exploring the data

Some summary statistics showing per state the total number of institutions, total amount awarded, total number of recipients.

```{r}
dat %>% 
  group_by(state) %>% 
  summarise(inst_total = length(unique(name)), award_total = sum(award, na.rm = TRUE), recipient_total = sum(recipient)) %>% 
  arrange(desc(award_total))

top20 <- dat %>% 
  group_by(state) %>% 
  summarise(award_total = sum(award, na.rm = TRUE)) %>%
  ungroup %>% 
  arrange(desc(award_total)) %>% 
  head(20)
```

## Growth of amount awarded over the years

The amount awarded peaked in 2010 and has fallen slowly over the years since.

```{r}
windowsFonts(Helvetica = "TT Arial")
dat %>%
  group_by(year) %>% 
  summarise(award_total_m = (sum(award, na.rm = TRUE) / 1000000), state = unique(state)) %>% 
  ggplot(aes(year(year), award_total_m)) +
  geom_line() +
  scale_y_log10(labels = scales::label_number(suffix = "B", big.mark = ",")) +
  labs(title = "Total amount awarded in billions of $") +
  bbc_style()
```

```{r}
windowsFonts(Helvetica = "TT Arial")
dat %>%
  group_by(year) %>% 
  summarise(award_recipient = (sum(award, na.rm = TRUE) / sum(recipient, na.rm = TRUE))) %>% 
  ggplot(aes(year(year), award_recipient)) +
  geom_line() +
  labs(title = "Average amount awarded per recipient") +
  bbc_style()
```


## amount received by state
Average amount received per institution for each state. (In terms of total amount received per state, CA stands out)

The state 'FM' stands out with the highest average amount received per institution. FM is the abbreviation for Micronesia. The sole reccipient in this territory is the College of Micronesia 
```{r}
dat %>%
  filter(state %in% top20$state) %>% 
  group_by(state) %>% 
  summarise(award_total = sum(award, na.rm = TRUE), inst_total = length(unique(name)),
            award_inst = (award_total / inst_total) / 1000) %>% 
  ggplot(aes(fct_reorder(state, award_inst), award_inst)) +
  geom_col() +
  scale_y_continuous(labels = scales::label_number(suffix = "M", big.mark = ",")) +
  coord_flip() +
  labs(x = "Amount awarded in million $",
       y = "State",
       title = "average amount per state's institution") +
  bbc_style()


```

## Change in amount received compared to previous year

```{r}
growth <- dat %>%
  group_by(state, year) %>% 
  summarise(award = sum(award, na.rm = TRUE)) %>%
  pivot_wider(names_from = year, values_from = award) %>% 
  ungroup()

for(i in 20:3) {
  growth[,i] = growth[,i] - growth[,i - 1]
  
}
growth[[2]] <- rep(0, length(growth[[2]]))
growth <- growth %>% pivot_longer(!state, names_to = "year", values_to = "difference") %>% 
  mutate(diff_b = difference / 1000000,
         positive = if_else(difference > 0, TRUE, FALSE),
         year = lubridate::year(year))
```

```{r}
growth_plot <- growth %>%
  filter(state %in% top20$state) %>% 
  ggplot(aes(state, diff_b, fill = positive)) +
  geom_col(show.legend = FALSE) +
  coord_flip()  +
  scale_y_continuous(labels = scales::label_number(suffix = "B", big.mark = ",")) +
  bbc_style() +
  transition_states(
    year,
    transition_length = 2,
    state_length = 1
  ) +
  ggtitle("Year: {closest_state}")
  animate(growth_plot, fps = 5)
```


## Top recipients for each year
```{r}
dat %>%
  group_by(year) %>% filter(award == max(award, na.rm = TRUE)) %>% 
  summarise(award = award, inst = name)

```

