---
title: "Tidy Tuesday - Week 36"
format: gfm

---
```{r, echo = FALSE}
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"), fig.width = 10, fig.heigth = 6)
source("../functions/ren_style.R")
```


```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
```


```{r, message = FALSE}
## Load Tidy Tuesday data
sets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/sets.csv.gz')

```

## "What happened with LEGOs? They used to be simple"
```{r, message = FALSE, warning = FALSE, fig.alt = "Line chart of number of LEGO sets released between 1949 and 2022"}
sets %>% 
  group_by(year) %>%
  filter(num_parts > 1) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(year, n)) +
  geom_line(size = 0.8) +
  labs(y = "Number of sets (2 or more parts)",
       x = "year",
       title = "LEGO popularity over the years") +
  ren_style()
```
As we can see LEGO has gained immense popularity over the years. The number of sets created each year has increased rapidly since the beginning. But let's explore the LEGO dataset further by answering the question "What happened with LEGOs? They used to be simple", a quote from one of my favorite TV-series Community. See the full quote here: "https://www.youtube.com/watch?v=yVCOAFKjaoY"

As a LEGO set, I defined any set consisting of 2 or more components. As we can see, the complexity of LEGO sets - defined here as the average number of components in a set - has steadily increased over the years. However, this is not the full story: by looking at the average, minimum and maximum number of components in sets over the years, we can see that complexity itself has not necessarily increased, but the *most complex* set contains more components over the years.

```{r, fig.alt = "Line chart of average number of components in LEGO sets released between 1949 and 2022"}

sets %>% 
  filter(num_parts > 1) %>% 
  group_by(year) %>% 
  summarise(mean = mean(num_parts)) %>% 
  ggplot(aes(year, mean)) +
  geom_line(size = 0.8, color = "#F8766D") +
  ren_style() +
  labs(title = "Average number of components in LEGO sets")
```
```{r, fig.alt = "Line chart of average, maximum and minimum components in LEGO sets, 1949 - 2022"}
sets %>%
  filter(num_parts > 1) %>% 
  group_by(year) %>% 
  summarise(mean = mean(num_parts), max = max(num_parts), min = min(num_parts)) %>% 
  ggplot(aes(year)) +
  geom_line(aes(y = mean, color = "Average"), size = 0.8) +
  geom_line(aes(y = max, color = "Most"), size = 0.8) +
  geom_line(aes(y = min, color = "Least"), size = 0.8) +
  ren_style() +
  labs(title = "Number of components in LEGO sets")
  
```



This can be further clarified by looking at the distribution of the number of components in LEGO sets for four years: 1949, 1968, 1986, 2004 and 2022. Here we can see that not only does the largest set become larger, the number of disproportionately large LEGO sets increases over the years as well.

```{r, fig.alt = "Boxplots of distribution of number of components in LEGO sets in 1949, 1968, 186, 2004 and 2022"}
sets %>% 
  filter(year %in% as.integer(summary(unique(sets$year)))) %>%
  filter(num_parts > 1) %>%
  mutate(year = as.factor(year)) %>% 
  group_by(year) %>% 
  ggplot(aes(x = year, y = num_parts, fill = year)) +
  geom_boxplot() +
  labs(y = "Number of components",
       x = "Year",
       title = "Distribution of number of set componenets") +
  ren_style() +
  theme(legend.position = "none")
```
```{r, message = FALSE}
ip_sets <- sets %>% 
  mutate(ip = case_when(
    stringr::str_detect(name, "Harry Potter") ==  TRUE ~ "Harry Potter",
    stringr::str_detect(name, "Star Wars") == TRUE ~ "Star Wars",
    stringr::str_detect(name, "Batman") == TRUE ~ "Batman"
  )
  )

ip_sets %>% 
  filter(!is.na(ip)) %>% 
  group_by(year, ip) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(year, n, color = ip)) +
  geom_line(size = 1) +
  ren_style() +
  labs(title = "Number of sets per year for different IPs")
```

Finally we can look at the number of LEGO sets for existing IPs such as Harry Potter and Star Wars. There is no clear growth in these kinds of sets, as the popularity of these sets is probably strongly linked to the release dates of the corresponding movies. I added Batman as a third IP, but these sets are not as popular apart from one peak year.

## Conclusion
Did LEGO sets get more complex? Yes, but no. They used to be simple and they still are. However the number and complexity of the *most complex* sets has increased. Finally, in terms of LEGO sets connected to existing IPs, there is not much to complain about except for 2017, the year of LEGO Batman :-)
